Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-10-07 doc/JFLA2022/paper.tex
improve (?) introduction of Section 2

This work is also an opportunity to 
 provide ``concrete'' examples of formalization and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc. It may be also used as a library on ordinal numbers, for instance for proving termination properties.

@@

\begin{theorem}
  Let $\mu$ be any ordinal strictly less than $\epsilon_0$.
  There is no function mapping hydras to the segment $[0,\mu)$ that could be used as a measure for  proving the termination of  all hydra battles.\label{thm3}
\end{theorem}

Considering measures applicable to \emph{all} battles allowed us to
focus on battles where the hydra can make an arbitrary number of copies at any time, which made our proof by contradiction artificially easier.
Unfortunately, the examples  most commonly shown in the litterature
(see for instance~\cite{KP82, bauer2008, BauerHydra}) 
assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
Without loss of generality, we assume the following restrictions:

@@

Using $H'$s equations as rewrite rules, we can study a realistic example. We take the hydra of figure~\ref{fig:start} and $i=0$ as initial configuration. 
By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
More generally, we prove that, for $\alpha\geq\omega^\omega$, 
the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.

>>>>

This work is also an opportunity to 
 provide concrete examples of formalization and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc. It may be also used as a library on ordinal numbers, for instance for proving termination properties.

@@

\begin{theorem}
  Let $\mu$ be any ordinal strictly less than $\epsilon_0$.
  There is no function mapping hydras to the segment $[0,\mu)$ that could be used as a measure for  proving the termination of  \emph{any} hydra battle.\label{thm3}
\end{theorem}

Our proof was based on the construction of a battle, where the
number of copies at each round is given by the elimination of an existential quantifier. So, it was mandatory to consider the class of \emph{all} hydra battles.

Unfortunately, the examples  most commonly shown in the litterature
(see for instance~\cite{KP82, bauer2008, BauerHydra}) 
assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
Since we are looking  for a minoration of the length of such battles,
we can work with the following hypotheses and invariants, 
without any loss of generality.

@@

Using $H'$s equations as rewrite rules, we can study a concrete example. We take the hydra of figure~\ref{fig:start} and $i=0$ as initial configuration. 
By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
More generally, we prove that, for $\alpha\geq\omega^\omega$, 
the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.
